Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over \cye{convex domains and} domains with shellable triangulations

### 2024-09-27 main.tex
first batch of changes

@@

\title{Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations}
\author{
    Th\'eophile Chaumont-Frelet\thanks{Affiliation of the First Author, \texttt{email1@example.com}} \and
    Martin Werner Licht\thanks{Affiliation of the Second Author, \texttt{martin.licht@epfl.ch}} \and
    Martin Vohral\'ik\thanks{Affiliation of the Third Author, \texttt{email3@example.com}}
}

@@

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\notice}[1]{\textcolor{red}{REMARK: #1}}

@@

\begin{abstract}
    We construct potentials for the gradient, the curl and the divergence operators, and more generally the exterior derivative, over domains with shellable triangulations. This class of triangulations includes local stars in dimension two and three. 
    The operator norms of our potentials satisfy explicitly computable bounds that only depend on the geometry. 
    We thus obtain computable upper bounds for Poincar\'e--Friedrichs inequalities and computable lower bounds for the eigenvalues of several vector Laplacians. 
    % 
    %     We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    %     This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    %     The main application that we envision are computable estimates over finite element patches. 
    %
    As a result with independent standing, we establish Poincar\'e--Friedrichs inequalities for the $L^{p}$ de~Rham complex over convex bounded domains. 
    This is achieved through bounding operator norms of Poincar\'e and Bogovski\u{\i} potential operators.
    We express all our main results in the calculus of differential forms and treat the gradient, curl, and divergence operator as its particular instances. 
\end{abstract}

@@

However, no such results are known over convex Lipschitz domains and for general Lebesgue exponents $1 \leq p \leq \infty$, which is why we rely on regularized potential operators as a remedy. % DONE: Theo : But -> However 
We follow closely the discussions by Costabel and McIntosh~\cite{costabel2010bogovskiui},
who study regularized Poincar\'e and Bogovski\u{\i}-type potential operators over domains star-shaped with respect to a ball. 
In comparison to their exposition, our variants of the regularized potential operators are simplified:
we only study them over convex domains and use simpler (that is, constant) weight functions. 
The resulting operator features generally lower regularity properties in comparison to the regularized potential in the literature;
however, their properties are sufficient for our purposes. We easily estimate their operator norms. 
We consider the latter outcome to be of independent interest. 
% DONE \todo{Theo : I believe Gallistl-Olkhovskiy also estimate the constant. They do star-shape and regularized stuff, so it's different, but we should keep that in mind and maybe reference them}.

@@

% Exterior calculus % DONE: MV recommends moving this down to finish the current line of thoughts 
\todo{What is the best place for this paragraph?}
Though we present our results in vector calculus, our main arguments are given in the formalism of exterior calculus~\cite{greub1967multilinear,lee2012smooth}. Exterior calculus is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite,gross2004electromagnetic,arnold2006finite,arnold2009geometric,arnold2010finite,demlow2014posteriori,licht2021local,arnold2021complexes}. Loosely speaking, exterior calculus is the calculus of antisymmetric tensor fields. The formalism is independent of the dimension and highlights underlying geometric structures common to the gradient, curl, and divergence operators.
%It also enables us to leverage on the rich literature of Sobolev spaces of differential forms.
For the purpose of our discussion, this formalism allows us to leverage results from a larger body of literature in differential geometry and functional analysis. 
% DONE: Theo : paragraph feels misplaced 
\\

@@

Whenever $\calT$ is a simplicial complex, 
the set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
% Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. In practice, we not always distinguish between points and singleton simplices. 
% Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
Similarly, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ refer to the vertices and the faces (that is, codimension one members) of this triangulation.\footnote{
% \begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics.~\cite{schrijver1998theory}
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra.~\cite{ziegler2012lectures}
% \end{remark}
}
% Given $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$,
% and we write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$.\footnote{\todo{MWL: It might be more economical to get rid of the supersimplex set, it is only used in this section to define patches.}}

@@

\todo{MWL: The different geometric quantities in this section should really be just what is needed later. Any shape measure should be locally measurable and close to what appears in the right-hand sides of the main results. Any inequalities here should be as tight as possible. }

@@

    \todo{Tighter estimate via geometry}

@@

\todo{What if $\ell = 0$ and $u$ is constant? The reconstruction is not complete for constants}

@@

    The operators discussed by Costabel and McIntosh are pseudodifferential operators of negative order, 
    because their averaging uses a smooth weight, and which proves boundedness of the operators between a variety of function spaces. 
    Explicit bounds for the higher-order seminorms of these pseudodifferential operators 
    have been contributed by Guzman and Salgado~\cite{guzman2021estimation}. % DONE: THEO That requires more comments 
    % Guzman and Salgado have estimated the operator norm. \todo{Describe Guzman and Salgado's contribution in more detail.} 
\end{remark}

@@

        We show that every function that is piecewise Lipschitz with respect to the triangulation $\calK$ must be Lipschitz over all of $|\calK|$. 
        We use a technique similar to Lemma~\ref{lemma:stars_are_shellable}. 
        After application of a piecewise linear Lipschitz mapping, 
        we can assume that $|\calK|$ is convex without loss of generality. 
        Any function piecewise Lipschitz with respect to $\calK$ must be Lipschitz over $|\calK|$. 
        \todo{not enough}

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_m$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ 
    and such that for each $T_{m} \in \calT$ in the shelling
    we have the following recursive estimate: 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
%         &
%         \leq  
%         C_{\PF,\Gamma,k,p}(\calT) 
%         \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
%         \Cfive{n}{k}(\calT)^{k+1} 
%         \diam(T)^{k} 
%         \| \cartan u \|_{L^{p}(U_{m-1})}
%         \\&\qquad
%         + 
%         C_{\PF,\Gamma,k,p}(\calT) 
%         \diam(T)^{k} 
%         \| \cartan u      \|_{L^{p}(T_{m})} 
%         \\&\qquad
%         + 
%         \Cdetsechs{n}{k}(\calT)^{-\frac 1 p} 
%         \Cfive{n}{k}(\calT)^{k} 
%         \| w \|_{L^{p}(U_{m-1})}
%         \\
        &
        \leq  
        C_{\PF,\Gamma,k,p}(\calT) 
        \diam(T)^{k} 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfive{n}{k}(\calT)^{k+1} 
            \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(U_{m-1})}
        \right)
        \\&\qquad
        + 
        \Cfive{n}{k}(\calT)^{k} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
    \todo{The notation $U_m$ should be settled in the definition of shellings.}
\end{theorem}

@@

\bibliographystyle{plain}
\bibliography{library}

>>>>

% Added by MV 
\usepackage{bm} % preferred version of bold symbols
\usepackage[normalem]{ulem} % for strikethrough (\sout) in collaborative writing
\usepackage{cancel} % for strikethrough in collaborative writing (math formulas)
\usepackage[textsize=tiny,color=blue!50!white]{todonotes}
\setlength{\marginparwidth}{2cm}

\newcommand\cye[1]{%
\protect\leavevmode
\begingroup
    \color{red!35!yellow}%
    #1%
\endgroup
}

@@

\title{Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over \cye{convex domains and} domains with shellable triangulations}
\author{
    Th\'eophile Chaumont-Frelet\thanks{Affiliation of the First Author, \texttt{email1@example.com}} \and
    Martin Werner Licht\thanks{Affiliation of the Second Author, \texttt{martin.licht@epfl.ch}} \and
    Martin Vohral\'ik\thanks{\cye{Inria, 48 rue Barrault, 75647 Paris, France \& CERMICS, Ecole des Ponts, 77455 Marne-la-Vall\'ee, France}, \texttt{\cye{martin.vohralik@inria.fr}}}
}

@@

\newcommand{\martin}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\notice}[1]{\textcolor{red}{REMARK: #1}}
\newcommand{\disk}{\cye{ball}}

@@

\begin{abstract}
    We construct potentials for the gradient, the curl\cye{,} and the divergence operators, and more generally the exterior derivative, over domains with shellable triangulations. 
    This class of triangulations \cye{in particular} includes local \cye{patches (stars)} in dimension two and three.
    The operator norms of our potentials satisfy explicitly computable bounds that only depend on the geometry. 
    We thus obtain computable upper bounds for Poincar\'e--Friedrichs inequalities and computable lower bounds for the eigenvalues of several vector Laplacians. 
    % 
    %     We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    %     This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    %     The main application that we envision are computable estimates over finite element patches. 
    %
    As a result with independent standing, we establish Poincar\'e--Friedrichs inequalities for the $L^{p}$ de~Rham complex over convex bounded domains. 
    This is achieved through \cye{a suitable construction of regularized} Poincar\'e and Bogovski\u{\i} potential operators \cye{where we can bound their operator norms}.
    \martin{We express all our main results in the calculus of differential forms and treat the gradient, curl, and divergence operator as its particular instances.} 
    \cye{Computational examples illustrate the theoretical findings.}
    \\

    We construct potentials for the gradient, the curl\cye{,} and the divergence operators, and more generally the exterior derivative, over domains with shellable triangulations. 
    This class of triangulations \cye{in particular} includes local \cye{patches (stars)} in dimension two and three.
    \cye{We derive} computable bounds \cye{on} the operator norms of our potentials that only depend on the geometry. We thus obtain computable upper bounds for \cye{constants in} Poincar\'e--Friedrichs inequalities \cye{as well as} computable lower bounds for the eigenvalues of several vector Laplacians.
    As a result with independent standing, we \cye{also} establish Poincar\'e--Friedrichs inequalities \cye{with computable constants} for the $L^{p}$ de~Rham complex over convex bounded domains.
    This is achieved through \cye{a suitable construction of regularized} Poincar\'e and Bogovski\u{\i} potential operators \cye{where we can bound their operator norms}.
    \cye{Computational examples illustrate the theoretical findings.}
\end{abstract}

\tableofcontents
\todo[inline]{MV: Added table of contents: for us, to see the structure of and the main terminology used in the paper}

@@

However, no such results are known over convex Lipschitz domains and for general Lebesgue exponents $1 \leq p \leq \infty$, which is why we rely on regularized potential operators as a remedy. % DONE: Theo : But -> However 
We follow closely the discussions by Costabel and McIntosh~\cite{costabel2010bogovskiui},
who study regularized Poincar\'e and Bogovski\u{\i}-type potential operators over domains star-shaped with respect to a ball. 
In comparison to their exposition, our variants of the regularized potential operators are simplified:
we only study them over convex domains and use simpler (that is, constant) weight functions. 
The resulting operator features generally lower regularity properties in comparison to the regularized potential in the literature;
however, their properties are sufficient for our purposes. We easily estimate their operator norms. 
We consider the latter outcome to be of independent interest. 
% DONE \martin{Theo : I believe Gallistl-Olkhovskiy also estimate the constant. They do star-shape and regularized stuff, so it's different, but we should keep that in mind and maybe reference them}.

@@

% Exterior calculus % DONE: MV recommends moving this down to finish the current line of thoughts 
\martin{What is the best place for this paragraph?}
Though we present our results in vector calculus, our main arguments are given in the formalism of exterior calculus~\cite{greub1967multilinear,lee2012smooth}. Exterior calculus is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite,gross2004electromagnetic,arnold2006finite,arnold2009geometric,arnold2010finite,demlow2014posteriori,licht2021local,arnold2021complexes}. Loosely speaking, exterior calculus is the calculus of antisymmetric tensor fields. The formalism is independent of the dimension and highlights underlying geometric structures common to the gradient, curl, and divergence operators.
%It also enables us to leverage on the rich literature of Sobolev spaces of differential forms.
For the purpose of our discussion, this formalism allows us to leverage results from a larger body of literature in differential geometry and functional analysis. 
% DONE: Theo : paragraph feels misplaced 
\\

@@

Whenever $\calT$ is a simplicial complex, 
the set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
% Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. In practice, we not always distinguish between points and singleton simplices. 
% Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
Similarly, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ refer to the vertices and the faces (that is, codimension one members) of this triangulation.\footnote{
% \begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics.~\cite{schrijver1998theory}
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra.~\cite{ziegler2012lectures}
% \end{remark}
}
% Given $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$,
% and we write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$.\footnote{\martin{MWL: It might be more economical to get rid of the supersimplex set, it is only used in this section to define patches.}}

@@

\martin{MWL: The different geometric quantities in this section should really be just what is needed later. Any shape measure should be locally measurable and close to what appears in the right-hand sides of the main results. Any inequalities here should be as tight as possible. }

@@

    \martin{Tighter estimate via geometry}

@@

\martin{What if $\ell = 0$ and $u$ is constant? The reconstruction is not complete for constants}

@@

    The operators discussed by Costabel and McIntosh are pseudodifferential operators of negative order, 
    because their averaging uses a smooth weight, and which proves boundedness of the operators between a variety of function spaces. 
    Explicit bounds for the higher-order seminorms of these pseudodifferential operators 
    have been contributed by Guzman and Salgado~\cite{guzman2021estimation}. % DONE: THEO That requires more comments 
    % Guzman and Salgado have estimated the operator norm. \martin{Describe Guzman and Salgado's contribution in more detail.} 
\end{remark}

@@

        We show that every function that is piecewise Lipschitz with respect to the triangulation $\calK$ must be Lipschitz over all of $|\calK|$. 
        We use a technique similar to Lemma~\ref{lemma:stars_are_shellable}. 
        After application of a piecewise linear Lipschitz mapping, 
        we can assume that $|\calK|$ is convex without loss of generality. 
        Any function piecewise Lipschitz with respect to $\calK$ must be Lipschitz over $|\calK|$. 
        \martin{not enough}

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_m$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ 
    and such that for each $T_{m} \in \calT$ in the shelling
    we have the following recursive estimate: 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
%         &
%         \leq  
%         C_{\PF,\Gamma,k,p}(\calT) 
%         \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
%         \Cfive{n}{k}(\calT)^{k+1} 
%         \diam(T)^{k} 
%         \| \cartan u \|_{L^{p}(U_{m-1})}
%         \\&\qquad
%         + 
%         C_{\PF,\Gamma,k,p}(\calT) 
%         \diam(T)^{k} 
%         \| \cartan u      \|_{L^{p}(T_{m})} 
%         \\&\qquad
%         + 
%         \Cdetsechs{n}{k}(\calT)^{-\frac 1 p} 
%         \Cfive{n}{k}(\calT)^{k} 
%         \| w \|_{L^{p}(U_{m-1})}
%         \\
        &
        \leq  
        C_{\PF,\Gamma,k,p}(\calT) 
        \diam(T)^{k} 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfive{n}{k}(\calT)^{k+1} 
            \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(U_{m-1})}
        \right)
        \\&\qquad
        + 
        \Cfive{n}{k}(\calT)^{k} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
    \martin{The notation $U_m$ should be settled in the definition of shellings.}
\end{theorem}

@@

\bibliographystyle{acm_mod}
\bibliography{library}
